---
title: "README"
author: "R. Mark Sharp"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---
[![Rdoc](http://www.rdocumentation.org/badges/version/roxygen2)](http://www.rdocumentation.org/packages/roxygen2)
[![Build Status](https://travis-ci.org/rmsharp/rmsutilityr.svg?branch=master)](https://travis-ci.org/rmsharp/crosssectiontimer)
[![codecov](https://codecov.io/gh/rmsharp/rmsutilityr/branch/master/graph/badge.svg)](https://codecov.io/gh/rmsharp/crosssectiontimer)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = 'H')
library("crosssectiontimer")
library(ggplot2)
library(kableExtra)
library(magrittr)
library(rmsutilityr)
library(stringi)

activity_file <- 
  system.file("extdata",
              "colony_management_defined_activities.csv", 
              package = "crosssectiontimer", lib.loc = NULL, 
              mustWork = FALSE)
value_stream_descriptor <- stri_c("husbandry and veterinary care of ",
                                  "colony animals")
actor <- "animal_caretaker" # can be any value in the actor column of the 
                            # activity file

# The product of n_actors, times_per_month, and n_months is the size. 
# The size should be about 250 for about 20 activities to get the precision
# illustrated herein
n_actors <- 20
times_per_month <- 5
n_months <- 12

iterations <- 5000          # 5000 seems more than adequate for stable results

activities <- get_defined_activities(activity_file)

```

# Introduction

A cross-sectional time on task estimation method is presented, compared to use
of time logging, illustrated with an example, and its accuracy and precision
demonstrated with a simulation. Finally, a proposed protocol for using this
time on task estimation technique is presented that discusses potential methods
to ensure representative time estimates are obtained.

This package contains the R functions used to create this document and the
simulation software.  The specific example, which is the 
`r value_stream_descriptor` is provided by setup section of the ".Rmd" source 
file
and a comma separated value (CSV) file with the actor, activity, and activity
descriptions. Users of this package can easily adapt the small customization
script and the task activities file to simulate the study they are considering
to estimate the precision they can expect for their specified task and sampling
design.

```{r child = "inst/extdata/installation.Rmd", ref = "installation"}
```

Find online documentation at 
[https://rmsharp.github.io/crosssectiontimer/](
https://rmsharp.github.io/crosssectiontimer/).

See the setup code chunk of this vignette to see how to customize this document
to examine results you can expect from your own time on task activity analysis.
You will need to provide the CSV file^[Comparable to
colony_management_defined_activities.csv] with your own actors^[Actors are used
as variable names so must follow rules for naming variables in R. They must
start with a letter, contain only letters, numerals, ".", and "_".],
activities, and 
activity descriptions in addition to providing your definitions for the setup
chunk definitions indicated below.

```{r key-value-pairs, eval = FALSE, echo = TRUE}
activity_file <- 
  system.file("extdata",
              "colony_management_defined_activities.csv", 
              package = "crosssectiontimer", lib.loc = NULL, 
              mustWork = FALSE)
value_stream_descriptor <- stri_c("husbandry and veterinary care of ",
                                  "colony animals")
actor <- "animal_caretaker" # can be any value in the actor column of the 
                            # activity file

# The product of n_actors, times_per_month, and n_months is the size. 
# The size should be about 250 for about 20 activities to get the precision
# illustrated herein
n_actors <- 20
times_per_month <- 5
n_months <- 12

iterations <- 5000          # 5000 seems more than adequate for stable results

```

